Mixture Models for Single-Cell Assays with Application to Vaccine Studies

### 2013-05-18 MIMOSA.tex
Reordered AUC on Figures for ICS data

\subsection{ICS}
In the ICS data we considered observations at the day 0 time point as true negatives, and observations at the day 182 time point as true positives (potentially underestimating the sensitivity of all methods considered here due to real non-responders at day 182). The MIMOSA model has higher sensitivity and specificity than method compared here for discriminating vaccine responders and non-responders  as shown by the ROC curves on Figure~\ref{fig:HVTN065}, panels A,C,E.
At an FDR between 10-20\%, MIMOSA would lead to about 20\% more true positives being detected. Our comparisons also show that ranking based on log-fold change alone is not reliable and should not be used.
In addition, MIMOSA gave estimates of the observed false discovery rate that are better or comparable to competing methods (Figure~\ref{fig:HVTN065}, panels B,D,F). Here we present the results based on IL2 and IFN-$\gamma$ alone and the subset IL2 and/or IFN-$\gamma$ that were used in the original study~\citep{Goepfert:2011ci}. These results are consistent for other cytokines and cytokine combinations (see Supplementary Figure 1).

>>>>

\subsection{ICS}
In the ICS data we considered observations from vaccinees and placebos at the day 0 time point and day 182 time point as true negatives, and observations from vaccinees at the day 182 time point as true positives. The MIMOSA model has higher sensitivity and specificity than method compared here for discriminating vaccine responders and non-responders  as shown by the ROC curves on Figure~\ref{fig:HVTN065}, panels A,C,E.
At an FDR between 10-20\%, MIMOSA would lead to about 20\% more true positives being detected.
In addition, MIMOSA gave estimates of the observed false discovery rate that are better or comparable to competing methods (Figure~\ref{fig:HVTN065}, panels B,D,F). Here we present the results based on IL2 and IFN-$\gamma$ alone and the subset IL2 and/or IFN-$\gamma$ that were used in the original study~\citep{Goepfert:2011ci}. These results are consistent for other cytokines and cytokine combinations (see Supplementary Figure 1).  We also evaluated,through simulation, what would happen to the ROC curves when these assumptions about true responders and non-responders did not hold (\textit{i.e.,} when not all day 182 vaccinees are true responders, Supplementary Figure 6), and found that we are potentially underestimating the sensitivity of all methods considered here due to real non-responders at day 182.
